---
title: "_00_by_region"
author: "Julian Kleindiek"
date: "7/23/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

# load packages
library(tidyverse)
library(kableExtra)
library(janitor)
library(maps)
library(geosphere)
library(data.table)
library(forecast)
library(tseries)
library(fpp)
library(TSA)

# set paths
path = "/Users/juliankleindiek/Desktop/UChicago/08_Quarter4/02_TimeSeries/Project"
filename = "_01_International_Report_Passengers.csv"
```


### Step 1. Read in data
```{r}
# set working directory
setwd(path)

# read in flight data 
df = fread(file.path(filename))

# save df as tibble
df = as_tibble(df)

# print available columns
names(df)

# head
head(df)
```

### Step 2. Data cleaning
```{r}
# to numeric
df$total = as.numeric(gsub(",", "", df$total))
df$scheduled = as.numeric(gsub(",", "", df$scheduled))
```


### Step 3. Data visualization
```{r}
# define regions
south_europe = c("PORTUGAL", "SPAIN", "MALTA", "CYPRUS", "FRANCE", "ITALY", "CROATIA", "GREECE")
southeast_asia = c("INDONESIA", "THAILAND", "MALAYSIA", "PHILIPPINES", "CAMBODIA") # "VIETNAM", "SINGAPUR", "LAOS"
south_america = c("BRAZIL", "ARGENTINA", "COLOMBIA", "PERU", "CHILE", "ECUADOR", "BOLIVIA")
oceania = c("AUSTRALIA", "NEW ZEALAND")
```

```{r}
# filter dat by given regions
df_south_europe = df %>% filter(fg_country %in% south_europe)
df_southeast_asia = df %>% filter(fg_country %in% southeast_asia)
df_south_america = df %>% filter(fg_country %in% south_america)
df_oceania = df %>% filter(fg_country %in% oceania)
```

```{r}
# aggregate by year
df_south_europe_y = aggregate(df_south_europe$total, by=list(cat1=df_south_europe$year), FUN=sum)
df_southeast_asia_y = aggregate(df_southeast_asia$total, by=list(cat1=df_southeast_asia$year), FUN=sum)
df_south_america_y= aggregate(df_south_america$total, by=list(cat1=df_south_america$year), FUN=sum)
df_oceania_y = aggregate(df_oceania$total, by=list(cat1=df_oceania$year), FUN=sum)

# create time series object
south_europe_ts_y = ts(df_south_europe_y$x, start=1990, end=2019, frequency = 1)
southeast_asia_ts_y = ts(df_southeast_asia_y$x, start=1990, end=2019, frequency = 1)
south_america_ts_y = ts(df_south_america_y$x, start=1990, end=2019, frequency = 1)
oceania_ts_y = ts(df_oceania_y$x, start=1990, end=2019, frequency = 1)

# plot
ts.plot(south_europe_ts_y, southeast_asia_ts_y, south_america_ts_y, oceania_ts_y,
        ylab="Number of passengers", xlab="Time", col=c("blue", "red", "green", "yellow"),
        main="International travel from the US by year")
legend("topleft", lty=c(1,1,1,1), col=c("blue", "red", "green", "yellow"),
       legend=c("South Europe", "Southeast Asia", "South America", "Oceania"))
```
\
- Note that 2019 data is not complete\


```{r}
# aggregate by month
df_south_europe_m = aggregate(df_south_europe$total, by=list(cat1=df_south_europe$year, 
                                                             cat2=df_south_europe$month), FUN=sum)
df_southeast_asia_m = aggregate(df_southeast_asia$total, by=list(cat1=df_southeast_asia$year, 
                                                             cat2=df_southeast_asia$month), FUN=sum)
df_south_america_m = aggregate(df_south_america$total, by=list(cat1=df_south_america$year, 
                                                             cat2=df_south_america$month), FUN=sum)
df_oceania_m = aggregate(df_oceania$total, by=list(cat1=df_oceania$year, 
                                                             cat2=df_oceania$month), FUN=sum)

# create time series object
south_europe_ts_m = ts(df_south_europe_m[order(df_south_europe_m$cat1),]$x, frequency = 12, start=c(1990,1))
southeast_asia_ts_m = ts(df_southeast_asia_m[order(df_southeast_asia_m$cat1),]$x, frequency = 12, start=c(1990,1))
south_america_ts_m = ts(df_south_america_m[order(df_south_america_m$cat1),]$x, frequency = 12, start=c(1990,1))
oceania_ts_m = ts(df_oceania_m[order(df_oceania_m$cat1),]$x, frequency = 12, start=c(1990,1))

# plot
ts.plot(south_europe_ts_m, southeast_asia_ts_m, south_america_ts_m, oceania_ts_m,
        ylab="Number of passengers", xlab="Time", col=c("blue", "red", "green", "yellow"),
        main="International travel from the US by month")
legend("topleft", lty=c(1,1,1,1), col=c("blue", "red", "green", "yellow"),
       legend=c("South Europe", "Southeast Asia", "South America", "Oceania"))
```


### Step 4. Data modeling
#### South Europe Monthly
```{r}
# plot the data
tsdisplay(south_europe_ts_m, main="Monthly passenger travel from the US to South Europe")
```
\
- The data is non-stationary having a positive trend\
- Further, it follows a seasonal pattern with a non-constant variance over time\

```{r}
# plot with box-cox transformation
lambda = BoxCox.lambda(south_europe_ts_m)
south_europe_ts_m_trans = BoxCox(south_europe_ts_m, lambda=lambda)
tsdisplay(south_europe_ts_m_trans)
```
\
- The Box-Cox transformation leads to a more constant variance over time\

```{r}
# test for stationarity of training data with KPSS test
kpss.test(south_europe_ts_m, null="Level")
```
\
- The null hypothesis for the KPSS test is that the data is stationary\
- Large p-values are indicative of stationarity and small p-values suggest non-stationarity\
- For this data, the p-value is 0.01 and hence the data is non-stationary at a 5% significance level\

```{r}
# deseasonalize data
south_europe_ts_m_deseasonal = diff(south_europe_ts_m, lag=12)

# plot results
tsdisplay(south_europe_ts_m_deseasonal)
```
\
- After deseasonalizing the data, we can observe no seasonality in the data anymore\
- In the PACF, there is a cutoff at lag 1 and at lag 12 allowing to draw the conclusion that p and P could both be 1\
- The ACF is now decaying, which it wasn’t prior to the seasonal differencing and which is a good sign\

```{r}
# select test and training data
train = window(south_europe_ts_m, c(2008,1), c(2018,6))
test = window(south_europe_ts_m, c(2018,7))
```

```{r}
# fit auto.arima() model
model_1 = auto.arima(train, stationary=FALSE, seasonal=TRUE)

# summary
summary(model_1)
```

```{r}
# investigate residuals
checkresiduals(model_1)
```
\
- The residuals for the plots of both models appear to be white noise\
- However, based on the Ljung-Box test, we cannot accept the null hypothesis of the data to be independent  since the p-value < 0.05\

#### South America Monthly
```{r}
# plot the data
tsdisplay(south_america_ts_m, main="Monthly passenger travel from the US to South America")
```
\
- The data is non-stationary having a positive trend\
- Further, it follows a seasonal pattern with a non-constant variance over time\

```{r}
# plot with box-cox transformation
lambda_2 = BoxCox.lambda(south_america_ts_m)
south_america_ts_m_trans = BoxCox(south_america_ts_m, lambda=lambda_2)
tsdisplay(south_america_ts_m_trans)
```
\
- The Box-Cox transformation leads to a more constant variance over time\

```{r}
# test for stationarity of training data with KPSS test
kpss.test(south_america_ts_m, null="Level")
```
\
- The null hypothesis for the KPSS test is that the data is stationary\
- Large p-values are indicative of stationarity and small p-values suggest non-stationarity\
- For this data, the p-value is 0.01 and hence the data is non-stationary at a 5% significance level\

```{r}
# deseasonalize data
south_america_ts_m_deseasonal = diff(south_america_ts_m, lag=12)

# plot results
tsdisplay(south_america_ts_m_deseasonal)
```
\
- After deseasonalizing the data, we can observe no seasonality in the data anymore\
- In the PACF, there is a cutoff at lag 1 and at lag 12 allowing to draw the conclusion that p and P could both be 1\
- The ACF is now decaying, which it wasn’t prior to the seasonal differencing and which is a good sign\

```{r}
# select test and training data
train_2 = window(south_america_ts_m, c(2008,1), c(2018,6))
test_2 = window(south_america_ts_m, c(2018,7))
```

```{r}
# fit auto.arima() model
model_2 = auto.arima(train_2, stationary=FALSE, seasonal=TRUE)

# summary
summary(model_2)
```

```{r}
# investigate residuals
checkresiduals(model_2)
```
\
- The residuals for the plots of both models appear to be white noise\
- Based on the Ljung-Box test, we can accept the null hypothesis of the data to be independent  since the p-value > 0.05\



### Step 5. Forecasting
#### South Europe Monthly
```{r}
# forecast the year 1980
h = 12
fc_1 = forecast(model_1, h=h)

# plot test data and forecasts of both models
autoplot(test, xlab = "Time", ylab = "Number of passengers",
         main="12-month forecast for South Europe Monthly with auto arima model", series="Test data") +
  autolayer(fc_1$mean, series="Forecast model_1")
```

```{r}
# accuracy of model_1
accuracy(fc_1, test)
```

```{r}
# forecast into the future
fc_2 = forecast(model_1, h=24, level=c(80, 95))

# plot the forecast
plot(fc_2, main="24-month forecast for South Europe Monthly with auto arima model")
```

#### South America Monthly
```{r}
# forecast the year 1980
h = 12
fc_3 = forecast(model_2, h=h)

# plot test data and forecasts of both models
autoplot(test_2, xlab = "Time", ylab = "Number of passengers",
         main="12-month forecast for South America Monthly with auto arima model", series="Test data") +
  autolayer(fc_3$mean, series="Forecast model_3")
```

```{r}
# accuracy of model_1
accuracy(fc_3, test)
```

```{r}
# forecast into the future
fc_4 = forecast(model_2, h=24, level=c(80, 95))

# plot the forecast
plot(fc_4, main="24-month forecast for South America Monthly with auto arima model")
```







### Open Questions:
- I still have a pattern in my residuals because of a second/offset seasonality; how can I integrate that into my model?







