---
title: "TS - Final Project - Modeling Test"
author: "Jerry Sha"
date: "8/3/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(fpp)
library(tseries)
library(forecast)
library(TSA)
library(zoo)

library(tidyverse)
library(xts)
```

# -------------------------------------------------------
# Load Data
# -------------------------------------------------------

```{r}
dataPath <- "E:/UChicago/4 - Summer 2020/Times Series Analysis/Final Project/"

flight_data <- read.csv(file = paste0(dataPath,"International_Report_Passengers.csv"),
                        header = TRUE)
```


```{r}
# Convert passenger counts from factors into numeric
# Note that commas are removed to avoid NAs
flight_data$scheduled <- as.numeric(as.character(gsub(",","", flight_data$scheduled)))
flight_data$charter <- as.numeric(as.character(gsub(",","", flight_data$charter)))
flight_data$total <- as.numeric(as.character(gsub(",","", flight_data$total)))

flight_data$data_dte <- as.Date(flight_data$data_dte , format = "%m/%d/%Y")
```

```{r}
head(flight_data, 3)
```

# -------------------------------------------------------
# Isolate carriers - Select samples for testing
# -------------------------------------------------------

```{r}
flight_data_AA <- flight_data %>% filter(carrier == "AA") # American Airlines
flight_data_DL <- flight_data %>% filter(carrier == "DL") # Delta
flight_data_UA <- flight_data %>% filter(carrier == "UA") # United Airlines
flight_data_AC <- flight_data %>% filter(carrier == "AC") # Air Canada
flight_data_BA <- flight_data %>% filter(carrier == "BA") # British Airways

# Additional carriers for manual merger adjustments
flight_data_US <- flight_data %>% filter(carrier == "US") # US Airways; merged with AA
flight_data_NW <- flight_data %>% filter(carrier == "NW") # Northwest; merged with DL
flight_data_CO <- flight_data %>% filter(carrier == "CO") # Continental; merged with UA
```


```{r}
flight_data %>% filter(carrier == "NW")
```



# -------------------------------------------------------
# Testing - American Airlines (AA)
# -------------------------------------------------------

```{r}
# Count American Airlines (AA) flights
flight_count_AA <- flight_data_AA %>%
  group_by(data_dte) %>%
  summarize(count = n()) %>%
  arrange(desc(desc(data_dte)))

flight_count_AA <- xts(flight_count_AA$count,flight_count_AA$data_dte)



# Count US Airways (US) flights
flight_count_US <- flight_data_US %>%
  group_by(data_dte) %>%
  summarize(count = n()) %>%
  arrange(desc(desc(data_dte)))

flight_count_US <- xts(flight_count_US$count,flight_count_US$data_dte)



# Combine AA and US time series to reflect merger
flight_count_AA_merged <- flight_count_AA + merge(flight_count_US, index(flight_count_AA), fill = 0)
head(flight_count_AA_merged)
```



```{r}
# Count American Airlines (AA) passengers
pass_count_AA <- flight_data_AA %>%
  group_by(data_dte) %>%
  summarize(total = sum(total)) %>%
  arrange(desc(desc(data_dte)))

pass_count_AA <- xts(pass_count_AA$total,pass_count_AA$data_dte)



# Count US Airways (US) passengers
pass_count_US <- flight_data_US %>%
  group_by(data_dte) %>%
  summarize(total = sum(total)) %>%
  arrange(desc(desc(data_dte)))

pass_count_US <- xts(pass_count_US$total,pass_count_US$data_dte)



# Combine AA and US time series to reflect merger
pass_count_AA_merged <- pass_count_AA + merge(pass_count_US, index(pass_count_AA), fill = 0)
head(pass_count_AA_merged)
```

```{r}
plot(flight_count_AA)
plot(pass_count_AA)
```

```{r}
plot(flight_count_AA_merged)
plot(pass_count_AA_merged)
```

+ United Airlines (UA) merged with US Airways (US) in 2013

```{r}
pass_count_AA_merged %>% 
  BoxCox(lambda = "auto") %>%
  plot(main = "Passenger Count - AA - Box-Cox Transformed")
```




```{r}
training_AA <- pass_count_AA_merged["2008-01/2018-05"]
testing_AA <- pass_count_AA_merged["2018-06/"]
```


```{r}
(arima_AA <- auto.arima(ts(training_AA,
                           frequency = 12),
                        seasonal = TRUE,
                        lambda = "auto"))
```



```{r}
plot(forecast(arima_AA,
              h = 12),
     include = 50)
```


# -------------------------------------------------------
# Testing - Delta (DL)
# -------------------------------------------------------

```{r}
# Delta (DL) flight count
flight_count_DL <- flight_data_DL %>%
  group_by(data_dte) %>%
  summarize(count = n()) %>%
  arrange(desc(desc(data_dte)))

flight_count_DL <- xts(flight_count_DL$count,flight_count_DL$data_dte)



# Northwest (NW) flight count
flight_count_NW <- flight_data_NW %>%
  group_by(data_dte) %>%
  summarize(count = n()) %>%
  arrange(desc(desc(data_dte)))

flight_count_NW <- xts(flight_count_NW$count,flight_count_NW$data_dte)



# Delta-Northwest merger
flight_count_DL_merged <- flight_count_DL + merge(flight_count_NW, index(flight_count_DL), fill = 0)

```

```{r}
# Delta (DL) passenger count
pass_count_DL <- flight_data_DL %>%
  group_by(data_dte) %>%
  summarize(total = sum(total)) %>%
  arrange(desc(desc(data_dte)))

pass_count_DL <- xts(pass_count_DL$total,pass_count_DL$data_dte)



# Northwest (NW) passenger count
pass_count_NW <- flight_data_NW %>%
  group_by(data_dte) %>%
  summarize(total = sum(total)) %>%
  arrange(desc(desc(data_dte)))

pass_count_NW <- xts(pass_count_NW$total,pass_count_NW$data_dte)



# Delta-Northwest merger
pass_count_DL_merged <- pass_count_DL + merge(pass_count_NW, index(pass_count_DL), fill = 0)
```


```{r}
plot(flight_count_DL)
plot(pass_count_DL)
```


```{r}
plot(flight_count_DL_merged)
plot(pass_count_DL_merged)
```

+ Delta (DL) merged with Northwest Airlines (NW) in 2008


```{r}
pass_count_DL_merged %>% 
  BoxCox(lambda = "auto") %>%
  plot(main = "Passenger Count - DL - Box-Cox Transformed")
```

```{r}
training_DL <- pass_count_DL_merged["2008-01/2018-05"]
testing_DL <- pass_count_DL_merged["2018-06/"]
```

```{r}
(arima_DL <- auto.arima(ts(training_DL,
                           frequency = 12),
                        seasonal = TRUE,
                        lambda = "auto"))
```

```{r}
plot(forecast(arima_DL,
              h = 12),
     include = 50)
```

# -------------------------------------------------------
# Testing - United Airlines (UA)
# -------------------------------------------------------

```{r}
# United Airlines (UA) flight count
flight_count_UA <- flight_data_UA %>%
  group_by(data_dte) %>%
  summarize(count = n()) %>%
  arrange(desc(desc(data_dte)))

flight_count_UA <- xts(flight_count_UA$count,flight_count_UA$data_dte)



# Continental (CO) flight count
flight_count_CO <- flight_data_CO %>%
  group_by(data_dte) %>%
  summarize(count = n()) %>%
  arrange(desc(desc(data_dte)))

flight_count_CO <- xts(flight_count_CO$count,flight_count_CO$data_dte)



# United-Continental merger
flight_count_UA_merged <- flight_count_UA + merge(flight_count_CO, index(flight_count_UA), fill = 0)
```

```{r}
# United Airlines (UA) passenger count
pass_count_UA <- flight_data_UA %>%
  group_by(data_dte) %>%
  summarize(total = sum(total)) %>%
  arrange(desc(desc(data_dte)))

pass_count_UA <- xts(pass_count_UA$total,pass_count_UA$data_dte)



# Continental (CO) passenger count
pass_count_CO <- flight_data_CO %>%
  group_by(data_dte) %>%
  summarize(total = sum(total)) %>%
  arrange(desc(desc(data_dte)))

pass_count_CO <- xts(pass_count_CO$total,pass_count_CO$data_dte)



# United-Continental merger
pass_count_UA_merged <- pass_count_UA + merge(pass_count_CO, index(pass_count_UA), fill = 0)
```

```{r}
plot(flight_count_UA)
plot(pass_count_UA)
```

```{r}
plot(flight_count_UA_merged)
plot(pass_count_UA_merged)
```

+ Spike in data comes from 2010-2012 merger with Continental (CO)

```{r}
pass_count_UA_merged %>% 
  BoxCox(lambda = "auto") %>%
  plot(main = "Passenger Count - UA - Box-Cox Transformed")
```

```{r}
training_UA <- pass_count_UA_merged["2008-01/2018-05"]
testing_UA <- pass_count_UA_merged["2018-06/"]
```


```{r}
(arima_UA <- auto.arima(ts(training_UA,
                           frequency = 12),
                        seasonal = TRUE,
                        lambda = "auto"))
```

```{r}
plot(forecast(arima_UA,
              h = 12),
     include = 50)
```



# -------------------------------------------------------
# Testing - Air Canada (AC)
# -------------------------------------------------------

```{r}
flight_count_AC <- flight_data_AC %>%
  group_by(data_dte) %>%
  summarize(count = n()) %>%
  arrange(desc(desc(data_dte)))

flight_count_AC <- xts(flight_count_AC$count,flight_count_AC$data_dte)

head(flight_count_AC)
```

```{r}
pass_count_AC <- flight_data_AC %>%
  group_by(data_dte) %>%
  summarize(total = sum(total)) %>%
  arrange(desc(desc(data_dte)))

pass_count_AC <- xts(pass_count_AC$total,pass_count_AC$data_dte)

head(pass_count_AC)
```


```{r}
plot(flight_count_AC)
plot(pass_count_AC)
```

+ Air Canada acquired Canadian Airlines in 2000

```{r}
pass_count_AC %>% 
  BoxCox(lambda = "auto") %>%
  plot(main = "Passenger Count - AC - Box-Cox Transformed")
```

```{r}
training_AC <- pass_count_AC["2008-01/2018-05"]
testing_AC <- pass_count_AC["2018-06/"]
```


```{r}
(arima_AC <- auto.arima(ts(training_AC,
                           frequency = 12),
                        seasonal = TRUE,
                        lambda = "auto"))
```

```{r}
plot(forecast(arima_AC,
              h = 12),
     include = 50)
```




# -------------------------------------------------------
# Testing - British Airways (BA)
# -------------------------------------------------------

```{r}
flight_count_BA <- flight_data_BA %>%
  group_by(data_dte) %>%
  summarize(count = n()) %>%
  arrange(desc(desc(data_dte)))

flight_count_BA <- xts(flight_count_BA$count,flight_count_BA$data_dte)

head(flight_count_BA)
```

```{r}
pass_count_BA <- flight_data_BA %>%
  group_by(data_dte) %>%
  summarize(total = sum(total)) %>%
  arrange(desc(desc(data_dte)))

pass_count_BA <- xts(pass_count_BA$total,pass_count_BA$data_dte)

head(pass_count_BA)
```


```{r}
plot(flight_count_BA)
plot(pass_count_BA)
```

```{r}
pass_count_BA %>% 
  BoxCox(lambda = "auto") %>%
  plot(main = "Passenger Count - BA - Box-Cox Transformed")
```


```{r}
training_BA <- pass_count_BA["2008-01/2018-05"]
testing_BA <- pass_count_BA["2018-06/"]
```


```{r}
(arima_BA <- auto.arima(ts(training_BA,
                           frequency = 12),
                        seasonal = TRUE,
                        lambda = "auto"))
```

```{r}
plot(forecast(arima_BA,
              h = 12),
     include = 50)
```



























