---
title: "final_project_LJ"
author: "Lola Johnston"
date: "8/10/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r}
library(ggplot2)
library(tseries)
library(forecast)
library(fpp)
library(dplyr)
library(tidyverse)
library(ggfortify)

data<- read_csv(file = '/Users/lolajohnston/Documents/001_school/004_summer_2020/003_python/Datacamp/flight_eda_v2.csv')
head(data)
```

# Chicago to Cancun


### 1. EDA done on Tableau\
\

### 2. Data Cleaning\
\
```{r}

# check for missing values
sum(is.na(data)) # 316618 missing values

# filter data to ohare as the starting airport
data_ord <- filter(data, usg_apt =='ORD')
# filter data to show heathrow as the ending airport
data_l <- filter(data_ord, fg_apt =='LHR') # HEATHROW
# filter data to show cancun as the ending airport
data_c <- filter(data_ord, fg_apt == 'CUN') # CANCUN
####################################SHORT CUT TO SHOW CANCUN DATA ####################3
####################################SHORT CUT TO SHOW CANCUN DATA ####################3
####################################SHORT CUT TO SHOW CANCUN DATA ####################3
####################################SHORT CUT TO SHOW CANCUN DATA ####################3
data_l <- data_c
####################################SHORT CUT TO SHOW CANCUN DATA ####################3
####################################SHORT CUT TO SHOW CANCUN DATA ####################3
####################################SHORT CUT TO SHOW CANCUN DATA ####################3
####################################SHORT CUT TO SHOW CANCUN DATA ####################3
####################################SHORT CUT TO SHOW CANCUN DATA ####################3

# check the structure of the data
str(data_l)
is.ts(data_l)
head(data_l)

# select the columns you need
df1<-data.frame(data_l$data_dte, data_l$usg_apt, data_l$fg_apt, data_l$total)
names(df1) <- c('date', 'usg_apt', 'fg_apt', 'total')
head(df1)

# aggregate the dates
df1<- aggregate(df1["total"], by=df1["date"], sum)
# order by date
df1<- df1[order(as.Date(df1$date, format="%d/%m/%Y")),]
# turn into time series object
london <- ts(df1[,2], start = c(1990,1),frequency=12)
head(london)

# check that it is time series object
is.ts(london)

# double check ts object is the right cycle
cycle(london)

# check for missing values
sum(is.na(london))

```
\
### 3. visualizing the data\
\
```{r}

autoplot(london)
ggtsdisplay(london)

# season plot: pretty cool view of year over year by month
ggseasonplot(london) 

# polar seasonal plot: time axis is circular (vs horizontal) 
ggseasonplot(london, polar = TRUE)

# subseries mini time plots: comprises mini time plots for each season, the mean for each season is shown as a blue horizontal line.
ggsubseriesplot(london)

# lag plot: not sure how to read this but it shows the lags?
gglagplot(london)

# Acf plot
ggAcf(london)

# box plot
boxplot(london~cycle(london),xlab="Date", ylab = "Passenger" ,main ="Monthly Air Passengers Boxplot")

```

\
### 4. decomposition of data\
\
```{r}

decomposeAP <- decompose(london,"multiplicative")
autoplot(decomposeAP)



```
Note: In these decomposed plots we can again see the trend and seasonality and the random component shown as “remainder”.\
\



### 5. test stationarity\
\
```{r}

adf.test(london) 


```
\
the p-value is not less than .05, we don't have strong enough evidence against the null hypothesis, so we fail to reject the null hypothesis. It is not stationary clearly.\
\

### 6. Forecasting - Part 1 | Baseline Forecasts: Naive & Seasonal Naive\
\
```{r}

#########################
#### Naive Forecast #####
#########################

# there are no parameters for naive forecast so you don't need a train/test split
fc_naive_london<- naive(london, h=18) # forecasting 18 months ahead
autoplot(fc_naive_london)
summary(fc_naive_london)
# Check the residuals from the naive forecasts 
london  %>% naive() %>% checkresiduals()
# not white noise yet, still far from it

# this is not a very good model because it doesn't capture seasonality


#########################
# Seasonal Naive Forecast 
#########################

fc_snaive_london <- snaive(london, h = 18)
autoplot(fc_snaive_london)
summary(fc_snaive_london)
# Check the residuals from the seasonal naive forecasts
london %>% snaive() %>% checkresiduals() 
# not white noise yet, still decaying exponentially

# this is a better model because it does capture seasonality

```
\

### 6.2 Forecasting - Part 2 | linear models
- tslm
- lm
```{r}

###########################
# linear model ###########
###########################

#autoplot(train) + geom_smooth(method="lm")+ labs(x ="Date", y = "Passenger numbers", #title="Air Passengers") 


# This is not a good model to fit as it doesn’t capture the seasonality and multiplicative effects over time.

```
\
### 6.3 Forecasting - Part 3 | More complicted models time to seperate data into train/test\
Models: Arima, SES, ETS, HW\
Section includes: 
- Train/Test/Split \
- Differencing\
- Transformations\

```{r}

#########################
# train test split 
#########################

# train and test split, test starts in 2012 ((confirm with group))
train <- window(london, end = c(2011, 12)) 
test <- window(london,start = c(2012,1))

#########################
# differencing
#########################

# differencing = 1 
train_d1 <- diff(train, differences = 1)
autoplot(train_d1) + ggtitle("differences = 1")
adf.test(train_d1) # p-value is 0.01 which is less than 0.05 so we accept alternative hypothesis that it is stationary

# trying different tests
kpss.test(train_d1) # kpss test to check for stationarity
kpss.test(train_d1, null = "Trend") 
kpss.test(train_d1, null = "Level")  # p-value is 0.1 so stationary

# differencing = 2 ( just checking, no need to difference twice after looking at results)
train_d2 <- diff(train, differences = 2)
autoplot(train_d2) + ggtitle("differences = 2")
kpss.test(train_d2) # kpss test to check for stationarity
kpss.test(train_d2, null = "Trend") 
kpss.test(train_d2, null = "Level")  # p-value is > 0.1 so stationary

# visualizing the differenced training data
cbind("original" = train,
      "1st order diff"  = diff(train, 1),
      "2nd order diff" = diff(train, differences=2)) %>%
  autoplot(facets=TRUE) +
  xlab("Year") +
  ggtitle("Differencing comparison")


#############################
#  box cox transformations 
#############################

lambda <- BoxCox.lambda(train)
lambda  #0.4846784 needs transformation then?, practically square root transformation
t_train <- train %>% BoxCox(lambda = lambda) 

autoplot(t_train)
autoplot(train) + ggtitle("Box Cox Transformation")
kpss.test(t_train, null = "Trend") # not stationary bc p-value is < 0.05
kpss.test(t_train, null = "Level") # not stationary bc p-value is < 0.05


```



```{r}

###########################
# autoarima
###########################

arima_london <- auto.arima(train, seasonal = TRUE)
arima_london # ARIMA(2,1,1)(2,1,1)[12] AICc=7573.95 
ggtsdiag(arima_london) # looks pretty good

fc_arima_london <- forecast(arima_london, level = c(95), h = 90)

autoplot(fc_arima_london)+ labs(title="Auto Arima") 


#########################
# exponential smoothing 
########################

# Use ses() to forecast the next h = 90 
fc_ses_london <- ses(train, h = 90)

# Use summary() to see the model parameters
summary(fc_ses_london)

# Use autoplot() to plot the forecasts
autoplot(fc_ses_london)+ labs(title="Simple Exponential Smoothing") 

# Add the one-step forecasts for the training data to the plot
#autoplot(fc_ses_london) + autolayer(fitted(fc_ses_london))
### keep getting error when trying to use autolayer -_-


#############################################
#### holt winters method, seasonal and trend
#############################################

fc_hw_a<- hw(train, seasonal = "additive", h = 90)
fc_hw_m<- hw(train, seasonal = "multiplicative", h = 90)

autoplot(fc_hw_a)+ labs(title="Holt Winters - Additive") 
autoplot(fc_hw_m)+ labs(title="Holt Winters - Multiplicative") 

# Check if residuals look like white noise
checkresiduals(fc_hw_a) #i think additive looks better?
checkresiduals(fc_hw_m)

##################################################################
#### exponential smoothing with Errors Trend and Seasonality (ETS)
##################################################################

# Fit ETS model
fit_ets <- ets(train)
# ETS(A,N,A)

# Check residuals
checkresiduals(fit_ets)

# Plot forecasts
autoplot(forecast(fit_ets))+ labs(title="ETS") 


```

